---
title: "Worksheet 3A"
author: "Bea Joyce C. Baylon"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Using VECTORS
#  A. You need to produce a vector that contains the first 11 letters.

```{r}
first_11_letters <- LETTERS[1:11]
first_11_letters
```

# B. Produce a vector that contains the odd numbered letters.

```{r}
odd_number_letters <- LETTERS[seq(1, 26, by=2)]
odd_number_letters
```

#  C. Produce a vector that contains the vowels

```{r}
vowels <- LETTERS[c(1, 5, 9, 15, 21)]
vowels
```
# Base on the above vector letters

#  D. Produce a vector that contains the last 5 lowercase letters.

```{r}
last5_lowercase <- letters[22:26]
last5_lowercase
```

#1 E. Produce a vector that contains letters between 15 to 24 letters in lowercase.

```{r}
lowercase_15_to_24 <- letters[15:24]
lowercase_15_to_24
```

# 2. Create a vector(not a dataframe) with the average temperatures in April for Tugue-garao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City. The average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.
# a.What is the R code and its result for creating a character vector for the city/townof Tuguegarao City, Manila, Iloilo City, Tacloban,Samal Island, and Davao City? Name theobject as city. The names should follow the same order as in the instruction.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city
```

# b.The average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.Name the object as temp. Write the R code and its output. Numbers should also follow what is in the instruction.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```

# c.Create a dataframe to combine the city and the temp by using ‘data.frame(). What the R code and its result?
```{r}
df_citytemp <- data.frame(city, temp)
df_citytemp
```

# d.Associate the dataframe you have created in 2.(c) by naming the columns usingthe names() function. Change the column names by using names() function as City and Temperature. What is the R code and its result?
```{r}
names(df_citytemp) <- c("City", "Temperature")
df_citytemp
```

# e.Print the structure by using str() function. Describe the output.
```{r}
str(df_citytemp)
```

# f.From the answer in d, what is the content of row 3 and row 4 What is its R code and its output?
```{r}
df_citytemp[3:4,]
```

# g.From the answer in d, display the city with highest temperature and the city with the lowest temperature. What is its R code and its output?
```{r}
highest_temp_city <- df_citytemp$City[which.max(df_citytemp$Temperature)]
lowest_temp_city <- df_citytemp$City[which.min(df_citytemp$Temperature)]

highest_temp_city
lowest_temp_city
```

# 2.  Create a matrix of one to eight and eleven to fourteen with four columns and three rows.

# a.What will be the R code for the #2 question and its result?
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
mat
```

# b.Multiply the matrix by two. What is its R code and its result?
```{r}
mult <- mat * 2
mult
```

# c. What is the content of row 2? What is its R code?
```{r}
mat[2, ]
```

# d.What will be the R code if you want to display the column 3 and column 4 in row 1 and row 2? What is its output?
```{r}
mat[1:2, 3:4]
```

# e.What is the R code is you want to display only the columns in 2 and 3, row 3? What is its output?
```{r}
mat[3, 2:3]
```

# f.What is the R code is you want to display only the columns 4? What is its output?
```{r}
mat[ ,4]
```

# g.Name the rows as isa, dalawa, tatlo and columns as uno, dos, tres, quatro for the matrix that was created in b.‘. What is its R code and corresponding output?
```{r}
rownames(mat) <- c("isa", "dalawa", "tatlo")
colnames(mat) <- c("uno", "dos", "tres", "quatro")
mat
```

# h.From the original matrix you have created in a, reshape the matrix by assigning a new dimension with dim(). New dimensions should have 2 columns and 6 rows. What will be the R code and its output?
```{r}
dim(mat) <- c(6, 2)
mat
```

# 3. An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1
# a.Create an array for the above numeric values. Each values will be repeated twice What will be the R code if you are to create a three-dimensional array with 4 columns and 2 rows. What will be its output?
```{r}
arr1 <- rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), 2)
numarray <- array(arr1, dim = c(2, 4, 3))
numarray
```

# b.How many dimensions do your array have?
```{r} dim(numarray)
```

# c.Name the rows as lowercase letters and columns as uppercase letters starting from
the A. The array names should be “1st-Dimensional Array”, “2nd-Dimensional Array”, and
“3rd-Dimensional Array”. What will be the R codes and its output?
```{r}
rownames(numarray) <- c("a", "b")
colnames(numarray) <- c("A", "B", "C", "D")
dimnames(numarray)[[3]] <- c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array")

numarray
```

