---
title: "Worksheet 4B"
author: "Bea Joyce C. Baylon"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#USING LOOP FUNCTION

# 1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r}
vectorA <- c(1, 2, 3, 4, 5)

matrix5x5 <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  matrix5x5[i, ] <- vectorA
}

print(matrix5x5)

```
# 2.Print the string "*" using for() function.
```{r}
num_rows <- 5
for (i in 1:num_rows) {
  cat(rep('"*"', i), "\n")
}
```

# 3. Get an input from the user to print the Fibonacci sequence starting from the 1st inputup to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
start <- as.integer(readline(prompt = "Enter the starting number for the Fibonacci sequence: "))
a <- 0
b <- 1
repeat {
  fib <- a + b
  a <- b
  b <- fib
  if (fib >= start) {
    cat(fib, " ")
  }
  if (fib >= 500) {
    break
  }
}
```
# Using Basic Graphics (plot(),barplot(),pie(),hist())
# 4. Import the dataset as shown in Figure 1 you have created previously.

# a.What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}
library(readxl)
shoe_size <- read_excel("shoe_size.xlsx")
head(shoe_size)
```
# b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
male_subset <- subset(shoe_size, Gender == "M")
female_subset <- subset(shoe_size, Gender == "F")

num_male <- nrow(male_subset)
num_female <- nrow(female_subset)

cat("Number of observations in Male:", num_male, "\n")
cat("Number of observations in Female:", num_female, "\n")
```

# c. Create a graph for the number of males and females for Household Data. Use plot(),chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
gender_counts <- table(shoe_size$Gender)

barplot(gender_counts,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("green", "purple"),
        legend = rownames(gender_counts),
        beside = TRUE)
```
# 5. The monthly income of Dela Cruz family was spent on the following

# a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
```{r}
expenses <- c(60, 10, 5, 25)
labels <- c("Food", "Electricity", "Savings", "Miscellaneous")
colors <- c("lightblue", "lightgreen", "pink", "purple")

pie(expenses,
    labels = paste(labels, "\n", round((expenses / sum(expenses)) * 100, 1), "%", sep = ""),
    col = colors,
    main = "Monthly Income Distribution of Dela Cruz Family")

```






