---
title: "Worksheet 4A"
author: "Bea Joyce C. Baylon"
date: "2024-10-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1.
```{r}
sframe <- data.frame(

Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),

Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),

Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
sframe
```
# a.
# The data contains two sets of observations for shoe size, height, and gender.

# b.

```{r}
males <- sframe[sframe$Gender == "M", c("Shoe_size", "Height")]
females <- sframe[sframe$Gender == "F", c("Shoe_size", "Height")]

males
females
```
#c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
mean_shoe_size <- mean(sframe$Shoe_size)
mean_height <- mean(sframe$Height)

mean_shoe_size
mean_height
```
#d. Is there a relationship between shoe size and height? Why?
```{r}
correlation <- cor(sframe$Shoe_size, sframe$Height)
correlation
```

#2. Construct character vector months to a factor with factor() and assign the result tofactor_months_vector. Print out factor_months_vector and assert that R prints outthe factor levels below the actual values.
#Consider data consisting of the names of months:"March","April","January","November","January","September","October","September","November","August","January","November","November","February","May","August","July","December","August","August","September","November","February",April")
```{r}
months_vector <- c(
"March", "April", "January", "November", "January", "September", "October",
"September", "November", "August", "January", "November", "November", "February",
"May", "August", "July", "December", "August", "August", "September", "November",
"February", "April")
months_vector

factor_months_vector <- factor(months_vector)

factor_months_vector
```
#3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)

summary(factor_months_vector)
```
#4. Create a vector and factor for the table below.
```{r}
directions_vector <- c("East", "West", "North")
frequencies_vector <- c(1, 4, 3)

factor_data <- factor(directions_vector)

new_order_data <- factor(factor_data, levels = c("East", "West", "North"))

new_order_data
```





