---
title: "Worksheet 3B BAYLON"
author: "Bea Joyce C. Baylon"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1. Create a data frame using the table below.

#a. Write the codes
```{r}
respondents_df <- data.frame(
  Respondents = 1:20,
  Sex = c(2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2),
  Fathers_Occupation = c(1, 3, 3, 3, 1, 2, 3, 1, 1, 1, 3, 2, 1, 3, 3, 1, 3, 1, 2, 1),
  Persons_at_Home = c(5, 7, 3, 8, 5, 9, 6, 7, 8, 4, 7, 5, 4, 7, 8, 8, 3, 11, 7, 6),
  Siblings_at_School = c(6, 4, 4, 1, 2, 1, 5, 3, 1, 2, 3, 2, 5, 5, 2, 1, 2, 5, 3, 2),
  Types_of_Houses = c(1, 2, 3, 1, 1, 3, 3, 1, 2, 3, 2, 3, 2, 2, 3, 3, 3, 3, 3, 2)
)

print(respondents_df)

```
#b. Describe the data. Get the structure or the summary of the data
```{r}
str(respondents_df)

summary(respondents_df)

```
#c. Is the mean number of siblings attending is 5?
```{r}
mean_siblings <- mean(respondents_df$Siblings_at_School)
print(mean_siblings)

```
#d. Extract the 1st two rows and then all the columns using the subsetting functions.Write the codes and its output.
```{r}
first_two_rows <- respondents_df[1:2, ]
print(first_two_rows)

```
#Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
subset_rows_cols <- respondents_df[c(3, 5), c(2, 4)]
print(subset_rows_cols)

```
#f. Select the variable types of houses then store the vector that results as types_houses.Write the codes.
```{r}
types_houses <- respondents_df$Types_of_Houses
print(types_houses)

```
#g. Select only all Males respondent that their father occupation was farmer. Write the
codes and its output.
```{r}
male_farmers <- respondents_df[respondents_df$Sex == 1 & respondents_df$Fathers_Occupation == 1, ]
print(male_farmers)

#2. Write a R program to create an empty data frame.
```{r}
df <- data.frame(
  Ints = integer(),
  Doubles = double(),
  Characters = character(),
  Logicals = logical(),
  Factors = factor(),
  stringsAsFactors = FALSE
)

print("Structure of the empty dataframe:")
str(df)
```
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.
```{r}
female_siblings <- respondents_df[respondents_df$Sex == 2 & respondents_df$Siblings_at_School >= 5, ]
print(female_siblings)

```

```








